---
title: "Code for TBCI Paper"
author: "Marie-Therese Meye, Hannah Reitz"
date: "23 January 2022"
output: html_document
---

## Packages
```{r include=FALSE}
pacman::p_load(
  tidyverse,
  knitr,
  kableExtra,
  readr,
  haven,
  stargazer,
  ggplot2,
  tableone,
  psych, 
  performance,
  scales,
  MatchIt,
  optmatch,
  cobalt,
  lmtest,
  sandwich,
  rgenoud,
  DescTools,
  jtools,
  ggplot2,
  ggthemes
)

```


## Loading Afrobarometer data Kenya round 8
```{r}
# Set working directory
setwd("G:\\My Drive\\HWS21\\Theory_Building_and_Causal_Inference\\Paper\\Code")

# Load data
data <- haven::read_sav("ken_r8.data_.new_.final_.wtd_release.31mar21.sav")
data

```


## Data Preparation

```{r}
# Renaming variables
data <- data %>%
  rename(violence = Q17D,
         elec = Q24,
         gender = Q101,
         age = Q1,
         food = Q7A,
         water = Q7B,
         medi = Q7C,
         fuel = Q7D,
         cash = Q7E,
         educ = Q97,
         rural = URBRUR,
         ethnic = Q81,
         party = Q91B, 
         id = RESPNO
         )


# Selecting all needed variables
data <- data %>%
  dplyr::select(id, violence, elec, gender, age, food, water, medi, fuel, cash, educ, rural, ethnic, party)

```                


## Deleting cases with missing values

```{r}
# Saving all factor variables as factors
str(data$violence)
data$violence <- factor(data$violence, ordered = TRUE)

str(data$elec)
data$elec <- factor(data$elec, ordered = TRUE)

str(data$gender)
data$gender <- factor(data$gender)

str(data$rural)
data$rural <- factor(data$rural)

str(data$food)
data$food <- factor(data$food, ordered = TRUE)
summary(data$food)

str(data$water)
data$water <- factor(data$water, ordered = TRUE)
summary(data$water)

str(data$medi)
data$medi <- factor(data$medi, ordered = TRUE)
summary(data$medi)

str(data$fuel)
data$fuel <- factor(data$fuel, ordered = TRUE)
summary(data$fuel)

str(data$cash)
data$cash <- factor(data$cash, ordered = TRUE)
summary(data$cash)

str(data$educ)
data$educ <- factor(data$educ, ordered = TRUE)
summary(data$educ)

str(data$ethnic)
data$ethnic <- factor(data$ethnic)
summary(data$ethnic)

str(data$party)
data$party <- factor(data$party)
summary(data$party)

  
summary(data)


# Setting missings

table(data$violence)
data$violence[data$violence == 9] <- NA
data$violence[data$violence == 8] <- NA

table(data$elec)
data$elec[data$elec == 5] <- NA
data$elec[data$elec == 9] <- NA

table(data$age)
data$age[data$age == -1] <- NA
data$age[data$age == 998] <- NA
data$age[data$age == 999] <- NA


table(data$water)
data$water[data$water == 8] <- NA

table(data$medi)
data$medi[data$medi == 9] <- NA

table(data$fuel)
data$fuel[data$fuel == 8] <- NA
data$fuel[data$fuel == 9] <- NA

table(data$cash)
data$cash[data$cash == 9] <- NA

table(data$educ)
data$educ[data$educ == 98] <- NA

table(data$party)
data$party[data$party == -1] <- NA
data$party[data$party == 9998] <- NA
data$party[data$party == 9999] <- NA

table(data$ethnic)
data$ethnic[data$ethnic == 9998] <- NA
data$ethnic[data$ethnic == 9999] <- NA


# Deleting all observations that have missing values on any of the variables
data <- na.omit(data)

summary(data)


# Dropping unnecessary levels of factor variables
data$violence <- droplevels(data$violence)
data$elec <- droplevels(data$elec)
data$water <- droplevels(data$water)
data$medi <- droplevels(data$medi)
data$fuel <- droplevels(data$fuel)
data$cash <- droplevels(data$cash)
data$water <- droplevels(data$water)
data$educ <- droplevels(data$educ)
data$ethnic <- droplevels(data$ethnic)
data$party <- droplevels(data$party)


summary(data)

```


## Recoding Variables

```{r}

# Binary violence variable / Treatment variable
table(data$violence)
data$treat <- ifelse(data$violence == '3', 0, 1)
data$treat <- as.factor(data$treat) # 0 = no fear of violence, 1 = fear of violence 
table(data$treat)

# Elections attitudes
data$elec <- recode_factor(data$elec, '1' = "0", '2' = "0", '3' = "1", '4' = "1") 

# Gender
data$gender <- ifelse(data$gender == '1', 0, 1)
data$gender <- as.factor(data$gender) # 0 = male, 1 = female
summary(data$gender)

# Education
table(data$educ)
data$educ <- recode_factor(data$educ, '0' = "0", '1' = "0", '2' = "1", '3' = "1", '4' = "2", '5' = "2", '6' = "3", '7' = "4", '8' = "4", '9' = "4") 
# 0 = No formal schooling, 1 = primary schooling, 2 = secondary schooling, 3 = post-secondary schooling, 4 = university
table(data$educ)

# Rural
data$rural <- recode_factor(data$rural, "1" = "0", "2" = "1") # 0 = urban, 1 = rural
table(data$rural)


# Ethnic tribe
table(data$ethnic)
data$ethnic <- recode_factor(data$ethnic, '302' = "0", '300' = "1", '301' = "2", '303' = "3", '304' = "4", '305' = "5", '306' = "5", '307' = "5", '308' = "5", '309' = "5", '310' = "5", '311' = "5", '312' = "5", '313' = "5", '314' = "5", '315' = "5", '316' = "5",'9990' = "5", '9995' = "5", '4' = "5") 
# 0 = Luhya, 1 = Kikuyu, 2 = Luo, 3 = Kamba, 4 = Kalenjin, 5 = Other
table(data$ethnic)

# Partisanship
table(data$party)
data$party <- recode_factor(data$party, '305' = "0", '300' = "1", '301' = "1", '302' = "1", '303' = "1", '304' = "1", '306' = "1", '307' = "1", '308' = "1", '309' = "1", '310' = "1", '311' = "1", '312' = "1", '313' = "1", '9995' = "1", '9997' = "2") 
# 0 = incumbent/government, 1 = oppostion, 2 = non-partisan
table(data$party)

summary(data)

```


## Lived poverty index reliability and creation

```{r}

# Converting lived poverty items to numeric
data$food <- as.numeric(data$food)
data$water <- as.numeric(data$water)
data$medi <- as.numeric(data$medi)
data$fuel <- as.numeric(data$fuel)
data$cash <- as.numeric(data$cash)

# Cronbach's alpha for lived poverty items
psych::alpha(subset(data, select = c(food, water, medi, fuel, cash)), check.keys = TRUE)

# Generating lived poverty index  
data <- data %>% 
  mutate(pov = (food + water + medi + fuel + cash) / 5)

summary(data)

# Deleting items for index
data <- subset(data, select = -c(food, water, medi, fuel, cash))

```


## Descriptive Statistics

```{r, echo = TRUE, include=TRUE}

attach(data)

myVars <- c("treat", "elec", "gender", "age", "educ", "rural", "pov", "ethnic", "party")

catVars <- c("treat", "elec", "gender", "educ", "rural", "ethnic", "party")

# Table of descriptive statistics
desctab <- CreateTableOne(vars = myVars, data = data, factorVars = catVars)
desctab
desc <- print(desctab, showAllLevels = TRUE, formatOptions = list(big.mark = ","), noSpaces = TRUE, catDigits = 2)

# Exporting table to Excel
write.csv(desc, file = "desc.csv")

```


## Regression without prior matching

```{r}
# Bivariate logistic regression of treatment on outcome
regwm1 <- glm(elec ~ treat, family=binomial(link="logit"), data = data)
summary(regwm1)
stargazer(regwm1, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Odd ratios
exp(coefficients(regwm1))


# Logistic regression with covariates
regwm2 <- glm(elec ~ treat + gender + age + educ + rural + pov + ethnic + party, family=binomial(link="logit"), data = data)
summary(regwm2)
stargazer(regwm2, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Odd ratios
exp(coefficients(regwm2))

```


## Exporting regression table

```{r}
stargazer(regwm1, regwm2, 
          align = TRUE, 
          title = "Results of logistic regressions estimating the effect of fear of electoral violence on attitudes towards democratic elections",
          dep.var.labels=c("Attitudes towards deomcratic elections","Attitudes towards deomcratic elections"),
          covariate.labels=c("Fear of electoral violence", "Female","Age","Primary schooling", "Secondary schooling", "Post-secondary schooling", "University schooling", "Rural", "Poverty level", "Kikuyu", "Luo",  "Kamba", "Kalenjin", "Other", "Opposition", "Non-partisan"),
          no.space=TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001),
          type = "latex")

```



## Generating and exporting coefficient plot

```{r}

coefplot1 <- plot_summs(regwm1, regwm2, scale = TRUE, colors = c("red", "blue"), 
           panel.border = "black",
           coefs = c("Fear of electoral viol." = "treat",
                     "Female" = "gender", "Age" = "age",
                     "Educ.: Primary" = "educ1", "Educ.: Secondary" = "educ2",
                     "Educ.: Post-secondary" = "educ3", "Educ.: University" = "educ4", 
                     "Rural" = "rural", "Poverty level" = "pov",
                     "Ethnic: Kikuyu" = "ethnic1", "Ethnic: Luo" = "ethnic2",
                     "Ethnic: Kamba" = "ethnic3", "Ethnic: Kalenjin" = "ethnic4", 
                     "Ethnic: Other" = "ethnic5", "Partisans.: Opposition" = "party1", 
                     "Partisans.: Non-partisan" = "party2")
           )

pdf(file = "Coefplot1.pdf",
    width = 7, 
    height = 4)

coefplot1 <- coefplot1 + theme_apa()  + labs(y = NULL) + theme(panel.grid.major.x=element_line(colour = "grey80"), axis.text.y = element_text(colour="black"), axis.text.x = element_text(colour="black")) + geom_vline(xintercept = 0, color = "black") + scale_x_continuous(breaks = seq(-1.5,0.5,by = 0.5)) + coord_cartesian(xlim = c(-1.5,1))

coefplot1

dev.off()

```


## Perform matching and selecting matching method

```{r}

# Selected matching method with the best balance
m_out1 <- matchit(
  treat ~gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "genetic",
  mahvars = ~ gender + age + educ + rural + pov + ethnic + party,
  caliper = c(.05, pov = 0.2),
  std.caliper = c(TRUE, FALSE),
  pop.size = 500,
  replace = TRUE,
  ratio = 1)
m_out1



# Other matching methods we tried
m_out2 <- MatchIt::matchit(
  treat ~ gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "full",
  distance = "mahalanobis",
  ratio = 1,
  replace = TRUE
)
m_out2

m_out3 <- MatchIt::matchit(
  treat ~ gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "nearest",
  distance = "mahalanobis",
  ratio = 1,
  replace = TRUE
)
m_out3


m_out4 <- MatchIt::matchit(
  treat ~ gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "genetic",
  distance = "mahalanobis",
  ratio = 1,
  replace = TRUE,
  pop.size = 100
)
m_out4


m_out5 <- MatchIt::matchit(
  treat ~ gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "nearest",
  distance = "logit",
  ratio = 1, 
  replace = TRUE
)
m_out5

m_out6 <- MatchIt::matchit(
  treat ~ gender + age + educ + rural + pov + ethnic + party,
  data = data,
  method = "optimal",
  distance = "mahalanobis",
  ratio = 1, 
)
m_out6

```

## Generating and exporting balance plot
```{r}
# Vector for variable names
names <- c(age = "Age",
           gender = "Female",
           educ_0 = "Educ.: No formal schooling",
           educ_1 = "Educ.: Primary",
           educ_2 = "Educ.: Secondary",
           educ_3 = "Educ.: Post-secondary",
           educ_4 = "Educ.: University",
           rural = "Rural",
           ethnic_0 = "Ethnic: Luhya",
           ethnic_1 = "Ethnic: Kikuyu",
           ethnic_2 = "Ethnic: Luo",
           ethnic_3 = "Ethnic: Kamba",
           ethnic_4 = "Ethnic: Kalenjin",
           ethnic_5 = "Ethnic: Other",
           party_0 = "Partisans.: Incumbent",
           party_1 = "Partisans.: Opposition",
           party_2 = "Partisans.: Non-partisan",
           pov = "Poverty level"
            )

pdf(file = "Balance.pdf", 
    width = 7,
    height = 4)

cobalt::love.plot(m_out1, 
                  var.order = "unadjusted",
                  drop.distance = TRUE,
                  #abs = TRUE,
                  var.names = names,
                  line = TRUE,
                  thresholds = c(m = .1),
                  binary = "std",
                  colors = c("red", "blue"),
                  sample.names = c("Unweighted", "Weighted")
                  )

dev.off()

```


## Check balance
```{r}

m_out1_sum <- summary(m_out1)
m_out1_sum

```


## Generating matched data set
```{r}

data_matched <- get_matches(m_out1)
head(data_matched)

```


## Regression with matched data
```{r}
# Saving weights as integers
data_matched$weights <- as.integer(data_matched$weights)

# Bivariate logistic regression with matched data
regm1 <- glm(elec ~ treat, family=quasibinomial(link="logit"),
data = data_matched,
weights = weights
)
stargazer(regm1, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Adjustment of standard errors to clustering in pairs and multiple use of control units
regclus1 <- lmtest::coeftest(regm1, vcov. = vcovCL, cluster = ~ subclass + id)
stargazer(regclus1, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Odd ratios
exp(coefficients(regclus1))



# Logistic regression with matched data and covariates
regm2 <- glm(elec ~ treat + gender + age + educ + rural + pov + ethnic + party, family=quasibinomial(link="logit"),
data = data_matched,
weights = weights
)
summary(regm2)
stargazer(regm2, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Adjustment of standard errors to clustering in pairs and multiple use of control units
regclus2 <- lmtest::coeftest(regm2, vcov. = vcovCL, cluster = ~ subclass + id)
stargazer(regclus2, type = "text", star.cutoffs = c(0.05, 0.01, 0.001))

# Odd ratios
exp(coefficients(regclus2))

```

## Exporting regression table

```{r}

stargazer(regclus1, regclus2, 
          align = TRUE, 
          title = "Results of logistic regressions estimating the effect of fear of electoral violence on attitudes towards democratic elections after matching on covariates",
          dep.var.labels=c("Attitudes towards deomcratic elections","Attitudes towards deomcratic elections"),
          covariate.labels=c("Fear of electoral violence", "Female","Age","Primary schooling", "Secondary schooling", "Post-secondary schooling", "University schooling", "Rural", "Poverty level", "Kikuyu", "Luo",  "Kamba", "Kalenjin", "Other", "Opposition", "Non-partisan"),
          no.space=TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001),
          type = "latex")

```


## Generating and exporting coefficient plot

```{r}
coefplot2 <- plot_summs(regclus1, regclus2, scale = TRUE, colors = c("red", "blue"), 
           panel.border = "black",
           coefs = c("Fear of electoral viol." = "treat1",
                     "Female" = "gender", "Age" = "age",
                     "Educ.: Primary" = "educ1", "Educ.: Secondary" = "educ2",
                     "Educ.: Post-secondary" = "educ3", "Educ.: University" = "educ4", 
                     "Rural" = "rural", "Poverty level" = "pov",
                     "Ethnic: Kikuyu" = "ethnic1", "Ethnic: Luo" = "ethnic2",
                     "Ethnic: Kamba" = "ethnic3", "Ethnic: Kalenjin" = "ethnic4", 
                     "Ethnic: Other" = "ethnic5", "Partisans.: Opposition" = "party1", 
                     "Partisans.: Non-partisan" = "party2"
           ))

coefplot2


pdf(file = "Coefplot2.pdf", 
    width = 7,
    height = 4)

coefplot2 <- coefplot2 + theme_apa()  + labs(y = NULL) + theme(panel.grid.major.x=element_line(colour = "grey80"), axis.text.y = element_text(colour="black"), axis.text.x = element_text(colour="black")) + geom_vline(xintercept = 0, color = "black")

coefplot2

dev.off()

```

